---
title: "SDIC Sunspot/ISN Prediction"
author: "David J Jackson"
date: "5/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(prophet)
library(forecast)
library(lubridate)
```

## A brief intro to Sunspot data and Prediction

The first thing we need to understand is that the (raw) data used to predict (in the plot below) the values of "y" are not sunspots or even mean sunspots but are daily mean Wolf(Rn) value. The Wolf(mean daily) values are calculate using the formula, Wolf = Groups *10 +Spots. Two examples:

1G 10 + 20S = 30 Wolf 2G 10 + 10S = 30 Wolf

```{r}
sidc <-fread("../db/SN_d_tot_V2.0.csv",sep=";")
colnames(sidc) <- c("Year","Month","Day", "Fdate","Spots", "Sd","Obs" ,"Defin"  )
sidc$Ymd <- as.Date(paste(sidc$Year, sidc$Month, sidc$Day, sep = "-"))
sidc1<- sidc %>% filter(Ymd>="1850-01-01") %>% select(Ymd,Spots)
colnames(sidc1) <- c("ds","y")
summary(sidc1)
```

## Sunspot Activity and Predictions: Jan 1, 1850 - July 31, 2030
```{r}
df <-sidc1 
m <- prophet(seasonality.mode="multiplicative")
m <- add_seasonality(m, name="cycle_11year", period=365.25 * 11,fourier.order=5)
m <- fit.prophet(m, df)
future <- make_future_dataframe(m,periods=4000,freq="day")
forecast <- predict(m, future)
plot(m, forecast) +ggtitle("SIDC Sunspot Prediction: Jan. 1850 - Oct. 2019") +ylab("Predicted Mean Wolf(Ra)") +xlab("Years" )
```
```{r}
df.pred <- forecast %>% select(ds,yhat,yhat_lower,yhat_upper) 
forecast <-as.data.table(forecast)
forecast$ds <-as.Date(forecast$ds)
```

```{r}
data <- df.pred %>% filter(ds >="2014-01-01")

ggplot(data=data,aes(x=ds,y=yhat)) +geom_line() +geom_smooth(method="loess")  + 
ggtitle("SIDC Current Min.: Jan 1, 2014 - Feb. 29,2020") +ylab("Means  Wolf(R)")
```
```{r}
a <- sidc[Ymd >="2008-12-01",]
a$ma <-ma(a$Spots,order=365)
ggplot(data=a,aes(x=Ymd,y=Spots)) + geom_line() + geom_smooth(method="loess") + 
geom_line(data=a,aes(x=Ymd,y=ma,col="MA"))
```
```{r}

```

